\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-12-16 archive.tex
3-panel figure

@@

\section{Introduction: Expertise}

Human learners acquire expertise in a wide range of domains: some of
us become experts in calculus, or cooking, or biology, music, tennis,
or software engineering, to name just a few, and every child develops
expertise in natural language, intuitive physics~\cite{Pete?},
intuitive psychology (theory-of-mind), and motor control.  This
picture contrasts sharply with the current state of machine
intelligence, where a machine is built to be an expert in a single
domain, like boardgames~\cite{alphaGo}, medical
diagnosis~\cite{mycin,CNNsForDiseaseDetection}, theorem
proving~\cite{eurisko}, or visual object recognition~\cite{alexNet}.
Thus an outstanding challenge in the long-term program of building
more humanlike machines is to develop an algorithm that, like people,
autonomously acquires expertise across many different kinds of
domains.

We contribute a model of the development of expertise that combines
two key ingredients.  First, an expert needs a sufficiently expressive
knowledge representation.  Following a long tradition in cognitive
science and AI, we represent knowledge as programs, which prior work
has used to represent expertise in recognition and generation of
handwriting and speech~\cite{lake2015human}, intuitive theories (of
kinship, taxonomy, etc.)~\cite{Ullman2012}, and natural language
grammar and
semantics~\cite{DBLP:journals/cogsr/SchmidK11,piantadosi2011learning}.
Second, atop this program representation, experts possess two kinds of
domain expertise.  They have at their disposal a powerful, yet
specialized, repertoire of concepts and abstractions: e.g., in
architecture, these are concepts like `arch' or `foundation'; in
software engineering these are libraries of code and domain-specific
languages (DSLs).  Here our model represents knowledge as programs,
and so we identify this kind of expertise with a DSL.  Experts also
have knowledge of when and how to deploy these domain-specific
concepts efficiently when solving new problems: e.g., mathematicians
efficiently search the space of proofs, intuiting which lemmas are
appropriate when; expert chefs intuit which compositions of
ingredients are likely tasty, before they actually start cooking.  For
our model, this aspect of expertise corresponds to the ability to
quickly assemble new, useful programs out of its DSL.

We integrate these ideas into a model called DreamCoder which acquires
expertise through a novel kind of wake/sleep or `dream' learning. The
model iterates through wake cycles -- where it solves problems by
writing programs -- and a pair of sleep cycles: a sleep cycle that
grows its DSL by replaying experiences from waking and consolidating
them into new abstractions, and a sleep cycle that improves its
knowledge of how to write programs by training a neural network on
replayed experiences as well as `dreams', or samples, from its DSL.

DreamCoder builds on multiple generations of AI research, going back
to the 1960's~\cite{solomonoff1964formal} when program-learning was
proposed as a paradigm for general AI.  Broadly speaking recent work
has either developed neural approaches for learning to efficiently
deploy a fixed
DSL~\cite{devlin2017robustfill,balog2016deepcoder,NGDS,spiral}, or
developed symbolic approaches for representing and searching through
spaces of
programs~\cite{gulwani2011automating,solar2008program,DBLP:books/daglib/0070933}.
We were motivated by approaches that learn or grow the
DSL~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning}.
Our goal with DreamCoder is to show that the combination of
neurally-guided search and DSL learning is a uniquely powerful way of
building systems that, like human learners, autonomously acquire the
expertise needed to navigate a new domain of problems.

\section{Introduction: AI/PL/dreaming}

@@

Our system is called DreamCoder because it acquires these two kinds of
expertise through a novel kind of wake/sleep or `dream'
learning~\cite{hinton1995wake}, iterating through a wake cycle --
where it solves problems by writing programs -- and a pair of sleep
cycles: a sleep cycle that grows the DSL by replaying experiences from
waking and consolidating them into new abstractions (modeled as
fragments of code), and a sleep cycle that improves its knowledge of
how to write programs by training a neural network based on replayed
experiences as well as `dreams', or samples, from the DSL.

DreamCoder builds on two lines of prior work.  One line of work
considers the problem of learning new concepts, abstractions, or
`options' from
experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
while the other line of work considers the problem of learning how to
deploy those concepts
efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
goal with DreamCoder is to show that the combination of these ideas is
uniquely powerful, and pushes us toward program-writing systems that,
like human learners, autonomously acquire the expertise needed to
solve a new class of problems.

\section{Introduction: NIPS}

Much of everyday human thinking and learning can be understood in
terms of program induction: constructing a procedure that maps inputs
to desired outputs, based on observing example input-output pairs.
People can induce programs flexibly across many different domains,
often from just one or a few examples.  For instance, if shown that a
text-editing program should map ``Jane Morris Goodall'' to
``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey'' to
``R. E. Leakey''; if instead the first input mapped to ``Dr. Jane'' or
``Goodall, Jane'', we might guess the latter should map to
``Dr. Richard'' or ``Leakey, Richard'', respectively.

The FlashFill system~\cite{gulwani2011automating} embedded in
Microsoft Excel solves problems such as these and is probably the best
known practical program-induction algorithm, but researchers in
programming languages and AI have had successes in many domains, such as handwriting recognition and
generation~\cite{lake2015human}, procedural
graphics~\cite{ellis2017learning,DBLP:journals/corr/abs-1804-01118}, question
answering~\cite{johnson2017clevr} and robot motion
planning~\cite{devlin2017neural}.  These systems
work in different ways, but most hinge upon a carefully engineered
\textbf{Domain Specific Language (DSL)}.  This is especially true for
systems like FlashFill that induce a wide range of programs
very quickly, in a few seconds or less.  DSLs constrain the search
over programs with strong prior knowledge in the form of a restricted
inventory of programming primitives finely tuned to the domain: for
text editing, these are operations like appending and
splitting on characters.

In this work, we consider the problem of building agents that learn to solve
program induction tasks, and also the problem of acquiring the prior knowledge
necessary to quickly solve these tasks in a new domain.  Representative problems
in three domains are shown in Table~\ref{initialExampleDSL}.  Our solution is an
algorithm that grows or boostraps a DSL while jointly training a neural network
to help write programs in the increasingly rich DSL.

\begin{tabular}{}
  \end{tabular}

@@

     \State \textbf{function} recognitionSample$'(Q, x, \text{parent}, \text{argumentIndex}, \mathcal{D}, \mathcal{E}, \tau)$:
     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, parent $\in \mathcal{D}\cup\left\{\text{start},\text{var} \right\}$, argumentIndex $\in \mathbb{N}$, environment $\mathcal{E}$, type $\tau$
     \State \textbf{Output:} a program whose type unifies with $\tau$
     \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
     \State var $\gets$ an unused variable name
     \State body $\sim$ recognitionSample$'(Q,x,\text{parent},\text{argumentIndex},\mathcal{D},\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$%\Comment{Recursively sample function body}
     \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
     %\Endif
     \Else\Comment{Build an application to give something w/ type $\tau$}
     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
     \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
     \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
       Q_{\text{parent},e,\text{argumentIndex}}(x)&\text{ if }e\in \mathcal{D}\\
Q_{\text{parent},\text{var},\text{argumentIndex}}(x)/|\text{variables}|&\text{ if }e\in \mathcal{E}
       \end{cases}$
     \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
     \State newParent$\gets\begin{cases}
     e &\text{ if }e\in \mathcal{D}\\
     \text{var}&\text{ if }e\in \mathcal{E}\end{cases}$
     \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
     %   \State unify$(\tau,\beta)$
     \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
     \State $a_k\sim\text{recognitionSample}'(Q,x,\text{newParent},k,\mathcal{D},\mathcal{E},\alpha_k)$
     \EndFor
     \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
     \EndIf
     %% \Statex
     %% \Statex\textbf{where:}
     %% \State yield$(\tau) = \begin{cases}
     %%   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
     %%   \tau   &\text{ otherwise.}
     %% \end{cases}$ \Comment{Final return type of $\tau$}
     %% \State  args$(\tau) = \begin{cases}
     %%   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
     %%   []   &\text{ otherwise.}
     %% \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
   \end{algorithmic}
 \end{algorithm}
 \begin{figure}
   \begin{tikzpicture}
     \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
     \node(e)[draw, align=center] at ([xshift=3cm]x.east) {Domain-specific encoder\\
       e.g., RNN or CNN};
     \node(m)[draw,align = center] at ([xshift=1.1cm]e.east) {MLP};%\\Output: $(|\mathcal{D}| + 2)\times(|\mathcal{D}| + 1)\times A$};
     \node[draw,align = center](s) at ([xshift=3cm]m.east) {Enumerate\\(Algorithm~\ref{recognitionSample})};
     \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Frontier\\$\mathcal{F}_x$};

>>>>

\newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
      \node[rectangle] at (1,1) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (I-\name) at (0,-\y) {};
      \foreach \name / \y in {1,...,3}
          \node[neuron] (H-\name) at (1,-\y-0.5) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (O-\name) at (2,-\y) {};
      \foreach \source in {1,...,4}
          \foreach \dest in {1,...,3}
              \draw [-latex] (I-\source) -- (H-\dest);
      \foreach \source in {1,...,3}
          \foreach \dest in {1,...,4}
              \draw [-latex] (H-\source) -- (O-\dest);
    \end{tikzpicture}}
\newcommand{\spiral}[2]{
  \draw[ultra thick,->] ([shift={#1}]-30:#2) arc [radius = #2, start angle = -30, end angle = 90];
  \draw[ultra thick,->] ([shift={#1}]-30:#2) arc [radius = #2, start angle = -30, end angle = 95];

      \draw[ultra thick,->] ([shift={#1}]90:#2) arc [radius = #2, start angle = 90, end angle = 210];
      \draw[ultra thick,->] ([shift={#1}]90:#2) arc [radius = #2, start angle = 90, end angle = 205];

      \draw[ultra thick,->] ([shift={#1}]210:#2) arc [radius = #2, start angle = 210, end angle = 340];
      \draw[ultra thick,->] ([shift={#1}]210:#2) arc [radius = #2, start angle = 210, end angle = 335];
}
\newcommand{\legend}{
  \begin{tikzpicture}
    \node at (0,0) (uses){is};
    \draw[->,red] ([xshift=-0.6cm]uses.west)  -- (uses.west);
    \node at ([xshift=0.4cm]uses.east) {\NeuralNetwork{0.15}};
    \draw[thin] (-1,-0.4) rectangle (1.2,0.4);
  \end{tikzpicture}
  }

@@

\begin{figure}
  \begin{tikzpicture}[line width=0.4mm]
    \draw (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
    \node at (5.5,1.25) {\textsc{\textbf{Wake}}};

    \begin{scope}[shift={(0.5,0.5)}]
      \node[align=right] at (0,0) (d){\textbf{DSL}\\
        $f_1$, $f_2$, ...};
      \node[align=center] at ([yshift = -2cm]d) (t){\textbf{Task}\\
        \footnotesize            \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
        \footnotesize    \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} };

      \node at ([xshift = 1.25cm]t.east) (nn){\NeuralNetwork{0.25}};
      \node[align = center, text width = 1cm] at ([yshift = 0.7cm,xshift=0cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
      \draw [red,-{>[scale=0.2]}] (t.east) -- ([xshift = -0.5cm]nn.west);

      \node[draw,rounded corners, align=center, inner sep = 10] at ([xshift = 4.2cm,yshift = 1cm]t.east) (s){Neurally-Guided\\ Enumerative Search};

      \draw [red,->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
      \draw [->,rounded corners,] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);

      \node[align=left] at ([xshift=3cm]s.east) (f) {\textbf{Programs:}\\
        \small    \code{($f_1$ $\ell$ ($\lambda$ (x) (> x 2)))}\\
        \small $\cdots\cdots\cdots$};
      \draw [->  ] (s.east) -- (f.west);

      \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
    \end{scope}
    \begin{scope}[shift={(9.4,-3.5)},scale=0.6,line width=0.05mm]
      \node[obs,scale=0.7] at (3.5,3) (dx){DSL};
      \node[latent,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
      \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
      \node[latent,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
      \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
      \draw [->] (zp1.south) -- (xp1.north);
      \draw [->] (dx.south) -- (zp1.north);
      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
      \node[latent,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
      \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
      \draw [->] (zp1.south) -- (xp1.north);
      \draw [->] (dx.south) -- (zp1.north);
      \draw [->] (dx.south) -- (zp.north);
      \draw [->] (zp.south) -- (xp.north);
      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
      \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
    \end{scope}

    \node at (0.5,-4.75) {\textbf{\textsc{Memory Consolidation}}};
    \draw (-2,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-2,-13) -- (-2,-5);
    \node at (11.5,-4.75) {\textbf{\textsc{Dreaming}}};
    \draw (14,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (14,-13) -- (14,-5);

    \begin{scope}[shift={(9,-4.5)}]
      \node(dreaming) at (2,-1) {\underline{Fantasies}};
      \node at ([yshift=-0.75cm,xshift=-1.5cm]dreaming.south) (d){DSL};
      \node at ([xshift=1.5cm]d.east) (p1){prog.};
      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
        \NeuralNetwork{0.17}};
      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
      \draw [dashed,red] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);

      \node(replay) at (2,-5) {\underline{Replays}};
      \node[align=center] at ([yshift=-0.75cm,xshift=-1.5cm]replay.south) (d){progs.\\for\\task};
      \node at ([xshift=1.5cm]d.east) (p1){prog.};
      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
        \NeuralNetwork{0.17}};
      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
      \draw [dashed,red] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);

      \begin{scope}[shift={(-3.3,-8)},scale=0.6,line width=0.05mm]
        \node[obs,scale=0.7] at (3.5,3) (dx){DSL};
        \node[obs,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
        \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
        \node[obs,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
        \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
        \draw [->] (zp1.south) -- (xp1.north);
        \draw [->] (dx.south) -- (zp1.north);
        \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
        \node[obs,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
        \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
        \draw [->] (zp1.south) -- (xp1.north);
        \draw [->] (dx.south) -- (zp1.north);
        \draw [->] (dx.south) -- (zp.north);
        \draw [->] (zp.south) -- (xp.north);
        \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
        \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
      \end{scope}

    %%}b memory consolidation
    \begin{scope}[shift={(-2,-4.5)}]
      \node[align=center] at (0.8,-1.2) (f1){progs. for\\task$_1$:\\\scode{(+ (car z) 1)}};
      \node[align=center] at ([xshift = 1cm]f1.east) (f2){progs. for\\task$_2$:\\\scode{(cons (+ 1 1))}};
      \node[align=center] at ([yshift = 0cm,xshift=1cm]f2.east) (f3){\textbf{DSL}\\\scode{cons}, \scode{+}, \\...};
      \node(c)[rectangle, rounded corners, dashed, draw, minimum width = 3.75cm, minimum height = 3cm, anchor = north] at ([yshift=-0.8cm]f2.south) {};
      \draw [-> ] (f1.south) -- (f1.south|-c.north);
      \draw [-> ] (f2.south) -- (f2.south|-c.north);
      \draw [-> ] (f3.south) -- (f3.south|-c.north);

      \node(d) at ([yshift = -1.8cm]c.south) {
        \begin{tikzpicture}[scale=0.9,line width=0.5mm]
          \node[align=center] at (0,0) {DSL w/ \code{(+ x 1)}:};
          \begin{scope}[shift={(0.6,-0.5)}]
            \node[pop3](p1) at (-1,-1) {\code{+}};
            \node[pop3](n1) at (0.6,-0.9) {\code{1}};
            \node[pop3](a) at (0,-1) {\code{ }};
            \draw[pop3] (0,0) -- (p1.north);
            \draw[pop3] (0,0) -- (n1.north);
            \draw[pop3] (-0.3,-0.3) -- (a.north);
          \end{scope}
      \end{tikzpicture}};
      \draw [-> ] (c.south) -- (d.north);

      \node at (c.center) {
        \begin{tikzpicture}[scale=0.9,line width=0.35mm]

          \begin{scope}[shift={(1,2)}]
            \node(l1) at (0,0) {};
            \node[color=pop3](p1) at (-1,-1) {\code{+}};
            \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
            %\node(x1) at (0,-1) {};
            \draw[color=pop3] (l1.south) -- (p1.north);
            \draw[color=pop3] (l1.south) -- (n1.north);
            \draw[color=pop3] (-0.5,-0.45) -- (0,-1);

      \begin{scope}[shift={(4,-8)},scale=0.6,line width=0.05mm]
        \node[latent,scale=0.7] at (3.5,3) (dx){DSL};
        \node[obs,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
        \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
        \node[obs,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
        \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
        \draw [->] (zp1.south) -- (xp1.north);
        \draw [->] (dx.south) -- (zp1.north);
        \node[obs,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
        \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
        \draw [->] (zp1.south) -- (xp1.north);
        \draw [->] (dx.south) -- (zp1.north);
        \draw [->] (dx.south) -- (zp.north);
        \draw [->] (zp.south) -- (xp.north);
      \end{scope}

    %% center spiral
    \begin{scope}[shift={(3.25,-7.9)},scale=0.8]    
      \spiral{(3.5,1)}{3.5}
      \node[latent,scale=1] at (3.5,3) (dx){DSL};
      \node[latent,scale=1] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
      \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
      \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
      \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
      \draw [->] (zp1.south) -- (xp1.north);
      \draw [->] (dx.south) -- (zp1.north);
      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);

      \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
      \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
      \draw [->] (zp1.south) -- (xp1.north);
      \draw [->] (dx.south) -- (zp1.north);
      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);

      \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
      \draw [->] (dx.south) -- (zp.north);
      \draw [->] (zp.south) -- (xp.north);

      \node at ([yshift=-0.6cm]xp.south) {\legend};

%% \section{Introduction: Expertise}

%% Human learners acquire expertise in a wide range of domains: some of
%% us become experts in calculus, or cooking, or biology, music, tennis,
%% or software engineering, to name just a few, and every child develops
%% expertise in natural language, intuitive physics~\cite{Pete?},
%% intuitive psychology (theory-of-mind), and motor control.  This
%% picture contrasts sharply with the current state of machine
%% intelligence, where a machine is built to be an expert in a single
%% domain, like boardgames~\cite{alphaGo}, medical
%% diagnosis~\cite{mycin,CNNsForDiseaseDetection}, theorem
%% proving~\cite{eurisko}, or visual object recognition~\cite{alexNet}.
%% Thus an outstanding challenge in the long-term program of building
%% more humanlike machines is to develop an algorithm that, like people,
%% autonomously acquires expertise across many different kinds of
%% domains.

%% We contribute a model of the development of expertise that combines
%% two key ingredients.  First, an expert needs a sufficiently expressive
%% knowledge representation.  Following a long tradition in cognitive
%% science and AI, we represent knowledge as programs, which prior work
%% has used to represent expertise in recognition and generation of
%% handwriting and speech~\cite{lake2015human}, intuitive theories (of
%% kinship, taxonomy, etc.)~\cite{Ullman2012}, and natural language
%% grammar and
%% semantics~\cite{DBLP:journals/cogsr/SchmidK11,piantadosi2011learning}.
%% Second, atop this program representation, experts possess two kinds of
%% domain expertise.  They have at their disposal a powerful, yet
%% specialized, repertoire of concepts and abstractions: e.g., in
%% architecture, these are concepts like `arch' or `foundation'; in
%% software engineering these are libraries of code and domain-specific
%% languages (DSLs).  Here our model represents knowledge as programs,
%% and so we identify this kind of expertise with a DSL.  Experts also
%% have knowledge of when and how to deploy these domain-specific
%% concepts efficiently when solving new problems: e.g., mathematicians
%% efficiently search the space of proofs, intuiting which lemmas are
%% appropriate when; expert chefs intuit which compositions of
%% ingredients are likely tasty, before they actually start cooking.  For
%% our model, this aspect of expertise corresponds to the ability to
%% quickly assemble new, useful programs out of its DSL.

%% We integrate these ideas into a model called DreamCoder which acquires
%% expertise through a novel kind of wake/sleep or `dream' learning. The
%% model iterates through wake cycles -- where it solves problems by
%% writing programs -- and a pair of sleep cycles: a sleep cycle that
%% grows its DSL by replaying experiences from waking and consolidating
%% them into new abstractions, and a sleep cycle that improves its
%% knowledge of how to write programs by training a neural network on
%% replayed experiences as well as `dreams', or samples, from its DSL.

%% DreamCoder builds on multiple generations of AI research, going back
%% to the 1960's~\cite{solomonoff1964formal} when program-learning was
%% proposed as a paradigm for general AI.  Broadly speaking recent work
%% has either developed neural approaches for learning to efficiently
%% deploy a fixed
%% DSL~\cite{devlin2017robustfill,balog2016deepcoder,NGDS,spiral}, or
%% developed symbolic approaches for representing and searching through
%% spaces of
%% programs~\cite{gulwani2011automating,solar2008program,DBLP:books/daglib/0070933}.
%% We were motivated by approaches that learn or grow the
%% DSL~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning}.
%% Our goal with DreamCoder is to show that the combination of
%% neurally-guided search and DSL learning is a uniquely powerful way of
%% building systems that, like human learners, autonomously acquire the
%% expertise needed to navigate a new domain of problems.

\section{Introduction}

@@

We call our system `DreamCoder' because it acquires these two kinds of
expertise through a novel kind of wake/sleep or `dream'
learning~\cite{hinton1995wake}, iterating through a wake cycle --
where it solves problems by writing programs -- and a pair of sleep
cycles, both of which are loosely biologically inspired by actual
human sleep.  The first sleep cycle, which we refer to as
\textbf{memory consolidation}, grows the DSL by replying experiences
from waking and consolidating them into new code abstractions.  This
cycle is inspired by the formation of abstractions during sleep memory
consolidation~\cite{DUDAI201520}.  The second sleep cycle, which we
refer to as \textbf{dreaming}, improves the agents knowledge of how to
write programs by training a neural network on replayed experiences as
well as `fantasies', or samples, from the DSL.  These two kinds of
dreams are inspired by the distinct episodic replay and hallucination
components of dream sleep~\cite{fosse2003dreaming}.

This dream-learning architecture unifies two lines of prior work.
One line of work
considers the problem of learning new concepts, abstractions, or
`options' from
experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
while the other line of work considers the problem of learning how to
deploy those concepts
efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
goal with DreamCoder is to show that the combination of these ideas is
uniquely powerful, and pushes us toward program-writing systems that,
like human learners, autonomously acquire the expertise needed to
navigate a new domain of problems.

\section{Deep Wake/Sleep Program Induction}

\system~takes as its goal to
acquire domain-specific expertise.
It gets this expertise from
solving a collection of
programming \textbf{tasks} -- typically about a hundred --
while jointly
authoring a library of code (the DSL)
and training a neural network

%% \section{Introduction: NIPS}

%% Much of everyday human thinking and learning can be understood in
%% terms of program induction: constructing a procedure that maps inputs
%% to desired outputs, based on observing example input-output pairs.
%% People can induce programs flexibly across many different domains,
%% often from just one or a few examples.  For instance, if shown that a
%% text-editing program should map ``Jane Morris Goodall'' to
%% ``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey'' to
%% ``R. E. Leakey''; if instead the first input mapped to ``Dr. Jane'' or
%% ``Goodall, Jane'', we might guess the latter should map to
%% ``Dr. Richard'' or ``Leakey, Richard'', respectively.

%% The FlashFill system~\cite{gulwani2011automating} embedded in
%% Microsoft Excel solves problems such as these and is probably the best
%% known practical program-induction algorithm, but researchers in
%% programming languages and AI have had successes in many domains, such as handwriting recognition and
%% generation~\cite{lake2015human}, procedural
%% graphics~\cite{ellis2017learning,DBLP:journals/corr/abs-1804-01118}, question
%% answering~\cite{johnson2017clevr} and robot motion
%% planning~\cite{devlin2017neural}.  These systems
%% work in different ways, but most hinge upon a carefully engineered
%% \textbf{Domain Specific Language (DSL)}.  This is especially true for
%% systems like FlashFill that induce a wide range of programs
%% very quickly, in a few seconds or less.  DSLs constrain the search
%% over programs with strong prior knowledge in the form of a restricted
%% inventory of programming primitives finely tuned to the domain: for
%% text editing, these are operations like appending and
%% splitting on characters.

%% In this work, we consider the problem of building agents that learn to solve
%% program induction tasks, and also the problem of acquiring the prior knowledge
%% necessary to quickly solve these tasks in a new domain.  Representative problems
%% in three domains are shown in Table~\ref{initialExampleDSL}.  Our solution is an
%% algorithm that grows or boostraps a DSL while jointly training a neural network
%% to help write programs in the increasingly rich DSL.

@@

     \State \textbf{function} recognitionSample$'(Q, x, \text{parent}, \text{argumentIndex}, \mathcal{D}, \mathcal{E}, \tau)$:
     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, parent $\in \mathcal{D}\cup\left\{\text{start},\text{var} \right\}$, argumentIndex $\in \mathbb{N}$, environment $\mathcal{E}$, type $\tau$
     \State \textbf{Output:} a program whose type unifies with $\tau$
     \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
     \State var $\gets$ an unused variable name
     \State body $\sim$ recognitionSample$'(Q,x,\text{parent},\text{argumentIndex},\mathcal{D},\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$%\Comment{Recursively sample function body}
     \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
     %\Endif
     \Else\Comment{Build an application to give something w/ type $\tau$}
     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
     \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
     \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
       Q_{\text{parent},e,\text{argumentIndex}}(x)&\text{ if }e\in \mathcal{D}\\
Q_{\text{parent},\text{var},\text{argumentIndex}}(x)/|\text{variables}|&\text{ if }e\in \mathcal{E}
       \end{cases}$
     \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
     \State newParent$\gets\begin{cases}
     e &\text{ if }e\in \mathcal{D}\\
     \text{var}&\text{ if }e\in \mathcal{E}\end{cases}$
     \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
     %   \State unify$(\tau,\beta)$
     \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
     \State $a_k\sim\text{recognitionSample}'(Q,x,\text{newParent},k,\mathcal{D},\mathcal{E},\alpha_k)$
     \EndFor
     \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
     \EndIf
     %% \Statex
     %% \Statex\textbf{where:}
     %% \State yield$(\tau) = \begin{cases}
     %%   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
     %%   \tau   &\text{ otherwise.}
     %% \end{cases}$ \Comment{Final return type of $\tau$}
     %% \State  args$(\tau) = \begin{cases}
     %%   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
     %%   []   &\text{ otherwise.}
     %% \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
   \end{algorithmic}
 \end{algorithm}
 \begin{figure}
   \begin{tikzpicture}[every node/.style={rounded corners,thick}]
     \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
     \node(e)[draw, align=center] at ([xshift=3cm]x.east) {Domain-specific encoder\\
       e.g., RNN or CNN};
     \node(m)[draw,align = center] at ([xshift=1.1cm]e.east) {MLP};%\\Output: $(|\mathcal{D}| + 2)\times(|\mathcal{D}| + 1)\times A$};
     \node[draw,align = center](s) at ([xshift=3cm]m.east) {Enumerate\\(Algorithm~\ref{recognitionSample})};
     \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Frontier\\$\mathcal{F}_x$};
